---
title: "p8105_hw2_sd3730"
author: "Stacey Dai"
output: github_document
---

# Question 1

This code is used to set up the RMarkdown file and load necessary packages.
```{r setup_q1, message = FALSE}
library(tidyverse)
library(dplyr)
```

This code is to read and clean the data, keep variables, and convert the entry variable from character to logical variable.
```{r cleanup_q1, message = FALSE, results = 'hide'}
transit_data = read.csv("./NYC_Transit_Subway_Entrance_And_Exit_Data.csv", 
                          na = c("NA", ".","")) |>
  janitor::clean_names() |>
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), entry,
         vending, entrance_type, ada) |>
    mutate(entry = ifelse(entry == "YES", TRUE, FALSE)) 
```

### Dataset Description

The NYC Transit dataset contains details about the stations, including station name and location, routes served by the stations, entrance and exits for each subway station in New York City. The dataset also contains information about vending machines in the station and ADA. 

The data cleaning steps involved reading the CSV file, cleaning the column names using janitor, selecting relevant columns, and converting the entry variable from character values ("YES"/"NO") to a logical variable (TRUE/FALSE). The resulting dataset has dimensions of `r nrow(transit_data)` rows and `r ncol(transit_data)` columns. The dataset is considered tidy as each variable forms a column, each observation forms a row, and each type of observational unit is a table.

How many distinct stations are there?
```{r stations_count, message = FALSE, results = 'hide'}
distinct_stations_count = transit_data |>
  distinct (station_name, line) |>
  nrow()
distinct_stations_count
```
There are `r distinct_stations_count` distinct subway stations.

How many stations are ADA compliant? 
```{r ada_count, message = FALSE, results = 'hide'}
ada_compliant_count = transit_data |>
  filter(ada == TRUE) |>
  nrow()
ada_compliant_count
```
There are `r ada_compliant_count` stations that are ADA compliant.

What proportion of station entrances / exits without vending allow entrance?
```{r no_vending, message = FALSE, results = 'hide'}
no_vending_proportion = transit_data |>
  filter(vending == "NO") |>
  summarise(proportion_entry_allowed = mean(entry, na.rm = TRUE)) |>
  pull(proportion_entry_allowed)

no_vending_proportion
```
Proportion of Entrances/Exits without Vending that Allow Entrance: `r no_vending_proportion` of entrances/exits without vending allow entrance.

Reformat data so that route number and route name are distinct variables.
```{r reformat, message = FALSE, results = 'hide'}
library(tidyr)

transit_data <- transit_data |>
  mutate(across(starts_with("route"), as.character)) 

transit_data_long <- transit_data |>
  pivot_longer(cols = starts_with("route"), 
               names_to = "route_number", 
               values_to = "route_name", 
               values_drop_na = TRUE)

distinct_a_train_stations <- transit_data_long |>
  filter(route_name == "A") |>
  distinct(station_name, line) |>
  nrow()

ada_compliant_a_train_count <- transit_data_long |>
  filter(route_name == "A", ada == TRUE) |>
  distinct(station_name, line) |>
  nrow()

distinct_a_train_stations
ada_compliant_a_train_count
```
There are `r distinct_a_train_stations` distinct stations that serve the A train. 
Of the stations that serve the A train, there are `r ada_compliant_a_train_count` compliant stations.


#Question 2

This code is used to set up the RMarkdown file and load necessary packages.
```{r setup_q2, message = FALSE}
library(readxl)
library(janitor)
library(dplyr)
```

Read and clean the Mr. Trash Wheel data
```{r mr_trash, message = FALSE, results = 'hide'}
mr_trash_wheel = read_excel("./202409 Trash Wheel Collection Data.xlsx", 
                              sheet = "Mr. Trash Wheel", 
                              skip = 1, 
                              col_names = TRUE) |>
    filter(!is.na(Dumpster)) |>
    select(Date, Dumpster, `Weight (tons)`, `Cigarette Butts`, `Sports Balls`) |>
    mutate(
        Trash_Wheel = "Mr. Trash Wheel",
        `Sports Balls` = as.integer(round(`Sports Balls`))
    )
```

Read and clean the Professor Trash Wheel data
```{r prof_trash, message = FALSE, results = 'hide'}
prof_trash_wheel = read_excel("./202409 Trash Wheel Collection Data.xlsx", 
                                sheet = "Professor Trash Wheel", 
                                skip = 1) |>
    filter(!is.na(Dumpster)) |>
    select(Date, Dumpster, `Weight (tons)`, `Cigarette Butts`) |>
    mutate(Trash_Wheel = "Professor Trash Wheel")
```

Read and clean the Gwynnda data
```{r gwynnda, message = FALSE, results = 'hide'}
gwynnda = read_excel("./202409 Trash Wheel Collection Data.xlsx", 
                      sheet = "Gwynnda Trash Wheel", 
                      skip = 1) |>
    filter(!is.na(Dumpster)) |>
    select(Date, Dumpster, `Weight (tons)`, `Cigarette Butts`) |>
    mutate(Trash_Wheel = "Gwynnda Trash Wheel")
```

```{r combine, message = FALSE, results = 'hide'}
combined_data = bind_rows(mr_trash_wheel, prof_trash_wheel, gwynnda)

head(combined_data)
```

```{r prof_total_weight, message = FALSE, results = 'hide'}
total_weight_professor = sum(combined_data$`Weight (tons)`[combined_data$Trash_Wheel == "Professor Trash Wheel"], na.rm = TRUE)
```

```{r total_cig_gwynnda, message = FALSE, results = 'hide'}
total_cigarette_butts_gwynnda_june <- sum(
  combined_data$`Cigarette Butts`[
    combined_data$Trash_Wheel == "Gwynnda Trash Wheel" & format(as.Date(combined_data$Date), "%Y-%m") == "2022-06"
  ],
  na.rm = TRUE
)

options(scipen = 999)
```

```{r total_obs, message = FALSE, results = 'hide'}
total_observations = nrow(combined_data)
```

The combined dataset comprises `r total_observations` observations from three trash wheel datasets: Mr. Trash Wheel, Professor Trash Wheel, and Gwynnda. Key variables include the date of collection, the dumpster used, the weight of trash collected in tons, the number of cigarette butts collected, and the number of sports balls collected. From the available data, the total weight of trash collected by Professor Trash Wheel is `r total_weight_professor` tons. Additionally, Gwynnda collected a total of `r total_cigarette_butts_gwynnda_june` cigarette butts in June 2022.


# Question 3

This code is used to set up the RMarkdown file and load necessary packages.
```{r setup_q3, message = FALSE, results = 'hide'}
library(dplyr)
library(tidyr)
library(readr)
library(janitor)
```

Read and clean the Great British Bake Off dataset
```{r gbb, message = FALSE, results = 'hide'}
bakers <- read_csv("./bakers.csv") |> clean_names()
bakes <- read_csv("./bakes.csv") |> clean_names()
results <- read_csv("./results.csv") |> clean_names()
viewers <- read_csv("./viewers.csv") |> clean_names()

glimpse(bakers)
glimpse(bakes)
glimpse(results)
glimpse(viewers)
```

Check for missing values.
```{r missing, message = FALSE, results = 'hide'}
sum(is.na(bakers))
sum(is.na(bakes))
sum(is.na(results))
sum(is.na(viewers))
```

Remove duplicates from the datasets.
```{r remove_dup, message = FALSE, results = 'hide'}
bakers <- bakers |> distinct()
bakes <- bakes |> distinct()
results <- results |> distinct()
viewers <- viewers |> distinct()
```


